---
title: "NYC Restaurant Inspection: Exploring the Data"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Setup

```{r Setup}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

## Loading libraries 
```{r}
library(tidyverse)
library(httr)
library(jsonlite)
library(plotly)
library(leaflet)
library(knitr)

theme_set(
  ggthemes::theme_fivethirtyeight() + theme(legend.position = "bottom")
  )

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.colour = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## 1.  Pulling in the dataset
```{r}
api_url = "https://data.cityofnewyork.us/resource/43nn-pn8j.csv"

rest_inspect = 
  GET(api_url, query = list("$limit" = 50000)) %>% 
  content("parsed")
```

## 2.  Data wrangling, cleaning up the dataset.
```{r}
rest_inspect = rest_inspect %>% 
  filter(grade %in% c("A", "B", "C"), boro != "0")
```

## 3.  Exploring dataset

### Grades

#### Counting how many restaurants by grade, by type.

```{r}
rest_inspect %>% 
  count(boro, grade) %>% 
  pivot_wider(names_from = grade, values_from = n) %>% 
  kable() 
```

#### Now considering different types of cuisine.

##### Italian/ Pizza Restaurants in NYC

```{r}
italian = c("Italian", "Italian/Pizza")

rest_inspect %>% 
  filter(str_detect(cuisine_description, italian)) %>% 
  count(boro, grade) %>% 
  pivot_wider(names_from = grade, values_from = n) %>% 
  kable()
```

##### French Restaurants in NYC

```{r}
rest_inspect %>% 
  filter(str_detect(cuisine_description, "French")) %>% 
  count(boro, grade) %>% 
  pivot_wider(names_from = grade, values_from = n) %>% 
  kable()
```

##### Mexican Restaurants in NYC
```{r}
rest_inspect %>% 
  filter(str_detect(cuisine_description, "Mexican")) %>% 
  count(boro, grade) %>% 
  pivot_wider(names_from = grade, values_from = n) %>% 
  kable()
```

##### Pan-Asian Restaurants in NYC

```{r}
asian = c("Asian", "Japanese", "Korean", "Chinese")

rest_grade =
  rest_inspect %>% 
  filter(str_detect(cuisine_description, asian)) %>% 
  count(boro, grade) %>% 
  pivot_wider(names_from = grade, values_from = n) %>% 
  kable()
```

### Critical flags

```{r}
rest_crit = 
  rest_inspect %>%
  count(boro, critical_flag) %>%   
  drop_na(critical_flag) %>% 
  pivot_wider(names_from = critical_flag, values_from = n) %>% 
  mutate(rest_ratio = N / (N+Y)) %>% 
  kable()
```

Year to Year Trends in New Restaurants
```{r}
rest_year_plot = 
  rest_inspect %>% 
  filter(str_detect(inspection_type, "Initial")) %>% 
  separate(grade_date, sep="-", into = c("grade_year", "grade_month", "grade_day")) %>%
  count(boro, grade_year) %>% 
  ggplot(aes(x = grade_year, y = n, color = boro)) + 
  geom_col((aes(fill = boro))) +
  labs(
    title = "New Restaurants in New York City",
    x = "Year",
    y = "Number of Restaurants")
rest_year_plot
```

### Restaurant Scores

#### Boxplot of Scores by Boro

```{r}
rest_inspect %>% 
  group_by(boro) %>% 
  filter(score %in% 1:30) %>% 
  ggplot(aes(x = boro, y = score, color = boro)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Scores",
    x = "Boro",
    y = "Score")
```



